---
title: "SCAG Data Task"
author: "Zack Crowley"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Set R chunk options:
## Set default options for R code chunks:
knitr::opts_chunk$set(
    echo = FALSE, 
    warning = FALSE, 
    message = FALSE 
    # fig.width = 6.5,
    # fig.height = 4
    )
# Prevents sci notation and sets the output of decimals to 4 (0.0000), suppresses warnings/info for tidyverse:
options(scipen = 999,
        tidyverse.quiet = TRUE,
        dplyr.summarise.inform = FALSE)

# Set here() location- working directory will be based on the project folder that this Rmd file is inside of currently:
here::i_am("SCAG_data_task_ZC.Rmd")

# Load libraries:
library(here)
library(fs)
library(rstudioapi) 
library(rlang)
library(rstatix)
library(janitor)
library(tidyverse)
library(readxl) # reading excel files
library(writexl) # writing excel files
library(data.table)
library(tidycensus) # API for US Census- can access ACS PUMS directly from API- key set in .Renviron- to see use: usethis::edit_r_environ()
library(survey)
library(srvyr) # Working with survey data

# # Get data dictionary as tibble from
# pums_vars_acs5 <- pums_variables %>%
#   filter(year == 2021, survey == "acs5")
# # person level
# pums_vars_acs5_person <- pums_variables %>%
#   filter(year == 2021, survey == "acs5", level == "person")
# # housing level:
# pums_vars_acs5_housing <- pums_variables %>%
#    filter(year == 2021, survey == "acs5", level == "housing")
# 
# # Write out data dicts for all vars, person and housing vars:
# write_csv(pums_vars_acs5, "data_dicts_ACS_PUMS/pums_vars_acs5.csv")
# write_csv(pums_vars_acs5_person, "data_dicts_ACS_PUMS/pums_vars_acs5_person.csv")
# write_csv(pums_vars_acs5_housing, "data_dicts_ACS_PUMS/pums_vars_acs5_housing.csv")
# 
# Read in data dicts from person and housign level data:
pums_vars_acs5 <- read_csv("data_dicts_ACS_PUMS/pums_vars_acs5.csv")
pums_vars_acs5_person <- read_csv("data_dicts_ACS_PUMS/pums_vars_acs5_person.csv")
pums_vars_acs5_housing <- read_csv("data_dicts_ACS_PUMS/pums_vars_acs5_housing.csv")

# Set up all file paths:
# file path for all raw data:
data_fp <- here('data')
# list.files(data_fp)

# file path for 2021 5-year ACS PUMS Data:
acs_2021_fp <- here(data_fp, '2017_2021_ACS_PUMS')
# list.files(acs_2021_fp)

# file path for 2021 person level data
person_2021_fp <- here(acs_2021_fp, "csv_pca", "psam_p06.csv")

# file path for 2021 household level data
hh_2021_fp <- here(acs_2021_fp, "csv_hca", "psam_h06.csv")

# Set up file path for 2022 5-year ACS PUMS Data:
acs_2022_fp <- here(data_fp, '2018_2022_ACS_PUMS')
# list.files(acs_2022_fp)

# file path for 2022 person level data
person_2022_fp <- here(acs_2022_fp, "csv_pca", "psam_p06.csv")

# file path for 2022 household level data
hh_2022_fp <- here(acs_2022_fp, "csv_hca", "psam_h06.csv")

# File paths and directories for clean data- 
# Creates a new folder called "clean_data" if it doesn't already exist in working directory:
if (!dir.exists("clean_data")) {dir.create("clean_data")}
# file path for clean_data
clean_data_fp <- here("clean_data")
# list.files(clean_data_fp)

# Clean 21 data
# Creates a new folder called "acs_data_21" (if it doesn't already exist in working directory)- this will be where the clean data for the 2021 ACS will be stored:
if (!dir.exists("acs_data_22")) {dir.create("acs_data_21")}
# file path for clean_data acs_data_21
acs_data_21_fp <- here(clean_data_fp, "acs_data_21")

# Clean 22 data
# Creates a new folder called "acs_data_22" (if it doesn't already exist in working directory)- this will be where the clean data for the 2021 ACS will be stored:
if (!dir.exists(here(clean_data_fp, 'acs_data_22'))) {dir.create(here(clean_data_fp, 'acs_data_22'))}
# file path for clean_data acs_data_22
acs_data_22_fp <- here(clean_data_fp, "acs_data_22")

# Clean ACS API data
# Creates a new folder called "acs_api" (if it doesn't already exist in working directory)- this will be where the clean data for the 2021 ACS will be stored:
if (!dir.exists(here(clean_data_fp, 'acs_api'))) {dir.create(here(clean_data_fp, 'acs_api'))}
# file path for clean_data acs_data_22
acs_api_fp <- here(clean_data_fp, "acs_api")

```

### Load in raw 2017-2021 ACS data from US Census microdata ACS PUMS data repo:

```{r load raw ACS 2021 data}
# Total list of vars to get from ACS PUMS
vars_list_pums <- c("SERIALNO", "PUMA", "ST", "SPORDER", "RELSHIPP", "PWGTP", "WGTP", "SEX", "AGEP", "HHT", "HISP", 
                    "RAC1P", "POVPIP", "NATIVITY", "DIS", "SCHL", "ESR", "JWMNP", 
                    "JWTRNS", "TEN","KIT", "PLM", "WAGP", "WKHP", 
                    "ADJINC","HINCP", "VEH","RMSP", "NP", "GRPIP", "OCPIP", "POVPIP")
       
### Housing level data raw:
# Variables to pull out of raw housing data:
housing_vars_list <- c("SERIALNO", "PUMA", "ST","WGTP", "HHT", "TEN", "KIT", "PLM", "ADJINC", "HINCP", "VEH", "RMSP", "NP", "GRPIP", "OCPIP")

# Read in the housing level data with file path named "person_2021_fp", select only vars from housing_vars_list, convert to tibble:
housing_data <- fread("data/csv_hca-Housing_CA/psam_h06.csv", select = housing_vars_list) %>% as_tibble()

### Person level data raw:
# Variables to pull out of raw housing data:
person_vars_list <- c("SERIALNO", "ST", "PUMA", "SPORDER", "RELSHIPP", "PWGTP", "SEX", "AGEP", "HISP", "RAC1P", "ENG", "NATIVITY", "DIS", "SCHL", "ESR", "JWMNP", "JWTRNS", "WAGP", "WKHP", "ADJINC", "POVPIP")
# Set up type list for read_csv:
# person_vars_list_types <- pums_vars_acs5 %>% 
#             distinct(var_code, data_type) %>% 
#             filter(var_code %in% person_vars_list) %>%
#             mutate(var_code =  factor(var_code, levels = person_vars_list)) %>%
#             arrange(var_code) %>% mutate(data_type = case_when(data_type == "chr" ~ "c",
#                                                                data_type == "num" ~ "d")) %>% 
#             tibble::deframe() %>% list()

# Read in the person level data with file path named "person_2021_fp":
person_data <- fread(here(person_2021_fp), select = person_vars_list) %>% as_tibble()
# person_data <- fread("data/csv_pca-Person_CA/psam_p06.csv", select = person_vars_list, 
#                      colClasses = list(numeric = c("ST", "PUMA", "SPORDER", "RELSHIPP", "PWGTP", "SEX", "AGEP", "HISP", "RAC1P", "ENG", "NATIVITY", 
#                                                    "DIS", "SCHL", "ESR", "JWMNP", "JWTRNS", "WAGP", "WKHP", "ADJINC", "POVPIP"))) %>% as_tibble()

# read_csv
# person_data <- read_csv("data/csv_pca-Person_CA/psam_p06.csv", col_select = person_vars_list, col_types = (list(SERIALNO = "c",
#                                                                                                                 ST       = "d",
#                                                                                                                 PUMA     = "d",
#                                                                                                                 RELSHIPP = "d",
#                                                                                                                 PWGTP    = "d",
#                                                                                                                 SEX      = "d",
#                                                                                                                 AGEP     = "d",
#                                                                                                                 HISP     = "d",
#                                                                                                                 RAC1P    = "d",
#                                                                                                                 ENG      = "c",
#                                                                                                                 NATIVITY = "d",
#                                                                                                                 DIS      = "d",
#                                                                                                                 SCHL     = "c",
#                                                                                                                 ESR      = "c",
#                                                                                                                 JWMNP    = "d",
#                                                                                                                 JWTRNS   = "c",
#                                                                                                                 WAGP     = "d",
#                                                                                                                 WKHP     = "d",
#                                                                                                                 ADJINC   = "c",
#                                                                                                                 POVPIP   = "d")), progress = show_progress()) 



# # Merge the person level data and the housing level data- using the variable `SERIALNO`; left_join on person_data- to only keep data in person-level data:
merged_data <- person_data %>% left_join(., housing_data, by = c("SERIALNO", "ST", "PUMA", "ADJINC"))

# Write out to .csv for the raw data from ACS Data call to folder named "clean_data/acs_data_21": 1826332 obs. with 31 variables
# write_csv(merged_data, "clean_data/acs_data_21/merged_acs_data.csv")

# SCAG County list- need to conver PUMA codes to county and then filter whole raw ACS PUMS data set to only these counties:
scag_county_list <- c("Imperial", 
                      "Los Angeles",
                      "Orange",
                      "Riverside",
                      "San Bernardino", 
                      "Ventura")

# Mutate/case_when to replicate SAS code above for creating "county" var- 
# Create couny variable with SCAG counties and filter to only those rows in SCAG Counties using scag_county_list from above:
merged_data_scag <- merged_data %>% mutate(county = case_when(PUMA == 2500 ~ "Imperial", # 25 prefix
                                      PUMA >= 3701 & PUMA <= 3799 ~ "Los Angeles", # 37 prefix
                                      PUMA >= 5901 & PUMA <= 5999 ~ "Orange", # 59 prefix
                                      PUMA >= 6501 & PUMA <= 6599 ~ "Riverside", # 65 prefix
                                      PUMA >= 7101 & PUMA <= 7199 ~ "San Bernardino", # 71 prefix
                                      PUMA >= 11101 & PUMA <= 11199 ~ "Ventura" # 111 prefix
                                                    )) %>% 
                            filter(county %in% scag_county_list) %>% relocate("county", .after = ST) 

# Write out to .csv for the raw data from ACS Data call to folder named "clean_data/acs_data_21": 879241 obs. (same as API call) with 33 variables
# write_csv(merged_data_scag, here(acs_data_21_fp, "merged_acs_data_scag.csv"))

# # check variables from the merged_data using str() and summary():
# str(merged_data)
```

### Load in raw 2017-2021 ACS data from US Census microdata ACS PUMS data repo:

```{r load raw ACS 2021 data}
# Total list of vars to get from ACS PUMS
vars_list_pums <- c("SERIALNO", "PUMA", "ST", "SPORDER", "RELSHIPP", "PWGTP", "WGTP", "SEX", "AGEP", "HHT", "HISP", 
                    "RAC1P", "POVPIP", "NATIVITY", "DIS", "SCHL", "ESR", "JWMNP", 
                    "JWTRNS", "TEN","KIT", "PLM", "WAGP", "WKHP", 
                    "ADJINC","HINCP", "VEH","RMSP", "NP", "GRPIP", "OCPIP", "POVPIP")
       
### Housing level data raw:
# Variables to pull out of raw housing data:
housing_vars_list <- c("SERIALNO", "PUMA", "ST","WGTP", "HHT", "TEN", "KIT", "PLM", "ADJINC", "HINCP", "VEH", "RMSP", "NP", "GRPIP", "OCPIP")

# Read in the housing level data, select only vars from housing_vars_list, convert to tibble:
housing_data <- fread("data/csv_hca-Housing_CA/psam_h06.csv", select = housing_vars_list) %>% as_tibble()

### Person level data raw:
# Variables to pull out of raw housing data:
person_vars_list <- c("SERIALNO", "ST", "PUMA", "SPORDER", "RELSHIPP", "PWGTP", "SEX", "AGEP", "HISP", "RAC1P", "ENG", "NATIVITY", "DIS", "SCHL", "ESR", "JWMNP", "JWTRNS", "WAGP", "WKHP", "ADJINC", "POVPIP")
# Set up type list for read_csv:
# person_vars_list_types <- pums_vars_acs5 %>% 
#             distinct(var_code, data_type) %>% 
#             filter(var_code %in% person_vars_list) %>%
#             mutate(var_code =  factor(var_code, levels = person_vars_list)) %>%
#             arrange(var_code) %>% mutate(data_type = case_when(data_type == "chr" ~ "c",
#                                                                data_type == "num" ~ "d")) %>% 
#             tibble::deframe() %>% list()

# Read in the person level data:
person_data <- fread("data/csv_pca-Person_CA/psam_p06.csv", select = person_vars_list) %>% as_tibble()
# person_data <- fread("data/csv_pca-Person_CA/psam_p06.csv", select = person_vars_list, 
#                      colClasses = list(numeric = c("ST", "PUMA", "SPORDER", "RELSHIPP", "PWGTP", "SEX", "AGEP", "HISP", "RAC1P", "ENG", "NATIVITY", 
#                                                    "DIS", "SCHL", "ESR", "JWMNP", "JWTRNS", "WAGP", "WKHP", "ADJINC", "POVPIP"))) %>% as_tibble()

# read_csv
# person_data <- read_csv("data/csv_pca-Person_CA/psam_p06.csv", col_select = person_vars_list, col_types = (list(SERIALNO = "c",
#                                                                                                                 ST       = "d",
#                                                                                                                 PUMA     = "d",
#                                                                                                                 RELSHIPP = "d",
#                                                                                                                 PWGTP    = "d",
#                                                                                                                 SEX      = "d",
#                                                                                                                 AGEP     = "d",
#                                                                                                                 HISP     = "d",
#                                                                                                                 RAC1P    = "d",
#                                                                                                                 ENG      = "c",
#                                                                                                                 NATIVITY = "d",
#                                                                                                                 DIS      = "d",
#                                                                                                                 SCHL     = "c",
#                                                                                                                 ESR      = "c",
#                                                                                                                 JWMNP    = "d",
#                                                                                                                 JWTRNS   = "c",
#                                                                                                                 WAGP     = "d",
#                                                                                                                 WKHP     = "d",
#                                                                                                                 ADJINC   = "c",
#                                                                                                                 POVPIP   = "d")), progress = show_progress()) 



# # Merge the person level data and the housing level data- using the variable `SERIALNO`; left_join on person_data- to only keep data in person-level data:
merged_data <- person_data %>% left_join(., housing_data, by = c("SERIALNO", "ST", "PUMA", "ADJINC"))

# Write out to .csv for the raw data from ACS Data call to folder named "clean_data/acs_data_21": 1826332 obs. with 31 variables
# write_csv(merged_data, here(acs_data_21_fp, "merged_acs_data_scag.csv"))

# SCAG County list- need to conver PUMA codes to county and then filter whole raw ACS PUMS data set to only these counties:
scag_county_list <- c("Imperial", 
                      "Los Angeles",
                      "Orange",
                      "Riverside",
                      "San Bernardino", 
                      "Ventura")

# Mutate/case_when to replicate SAS code above for creating "county" var- 
# Create couny variable with SCAG counties and filter to only those rows in SCAG Counties using scag_county_list from above:
merged_data_scag <- merged_data %>% mutate(county = case_when(PUMA == 2500 ~ "Imperial", # 25 prefix
                                      PUMA >= 3701 & PUMA <= 3799 ~ "Los Angeles", # 37 prefix
                                      PUMA >= 5901 & PUMA <= 5999 ~ "Orange", # 59 prefix
                                      PUMA >= 6501 & PUMA <= 6599 ~ "Riverside", # 65 prefix
                                      PUMA >= 7101 & PUMA <= 7199 ~ "San Bernardino", # 71 prefix
                                      PUMA >= 11101 & PUMA <= 11199 ~ "Ventura" # 111 prefix
                                                    )) %>% 
                            filter(county %in% scag_county_list) %>% relocate("county", .after = ST) 

# Write out to .csv for the raw data from ACS Data call to folder named "clean_data/acs_data_21": 879241 obs. (same as API call) with 33 variables
# write_csv(merged_data_scag, here(acs_data_21_fp, "merged_acs_data_scag.csv"))

# # check variables from the merged_data using str() and summary():
# str(merged_data)
```



### CA ACS PUMS Data from 2017-2021- 5 year ACS- Census API Call:
```{r tidycensus CA data 2021}
pums_vars_acs5_22 <- pums_variables %>% filter(year == 2022, survey == "acs5")
  
# List of vars to get from ACS PUMS
vars_list_pums <- c("PUMA","RELSHIPP","SEX", "AGEP", "HHT", "HISP", "RAC1P", "ENG", "NATIVITY", "DIS", "SCHL", "ESR", "JWMNP", "JWTRNS", 
                    "TEN", "KIT", "PLM", "WAGP", "WKHP", "ADJINC", "HINCP", "VEH",
                    "RMSP", "NP", "GRPIP", "OCPIP", "POVPIP")

# Return tibble of var_code and var_label for vars_list_pums, arrange by order of vars_list_pums:
pums_data_desc <- pums_vars_acs5 %>% 
            distinct(var_code, var_label, data_type, level) %>% 
            filter(var_code %in% vars_list_pums) %>%
            mutate(var_code =  factor(var_code, levels = vars_list_pums)) %>%
            arrange(var_code)

# List of vars and labels:
#   var_code  var_label                                                                                                                                   data_type   level  
#    <fct>    <chr>                                                                                                                                         <chr>     <chr>  
#  1 PUMA     Public use microdata area code (PUMA) based on 2010 Census definition (areas with population of 100,000 or more, use with ST for unique code) chr       NA     
#  2 RELSHIPP Relationship to reference person                                                                                                              chr       person 
#  3 SEX      Sex                                                                                                                                           chr       person 
#  4 AGEP     Age                                                                                                                                           num       person 
#  5 HHT      Household/family type                                                                                                                         chr       housing
#  6 HISP     Recoded detailed Hispanic origin                                                                                                              chr       person 
#  7 RAC1P    Recoded detailed race code                                                                                                                    chr       person 
#  8 ENG      Ability to speak English                                                                                                                      chr       person 
#  9 NATIVITY Nativity                                                                                                                                      chr       person 
# 10 DIS      Disability recode                                                                                                                             chr       person 
# 11 SCHL     Educational attainment                                                                                                                        chr       person 
# 12 ESR      Employment status recode                                                                                                                      chr       person 
# 13 JWMNP    Travel time to work                                                                                                                           num       person 
# 14 JWTRNS   Means of transportation to work                                                                                                               chr       person 
# 15 TEN      Tenure                                                                                                                                        chr       housing
# 16 KIT      Complete kitchen facilities                                                                                                                   chr       housing
# 17 PLM      Complete plumbing facilities                                                                                                                  chr       housing
# 18 WAGP     Wages or salary income past 12 months (use ADJINC to adjust WAGP to constant dollars)                                                         num       person 
# 19 WKHP     Usual hours worked per week past 12 months                                                                                                    num       person 
# 20 ADJINC   Adjustment factor for income and earnings dollar amounts (6 implied decimal places)                                                           chr       housing
# 21 HINCP    Household income (past 12 months, use ADJINC to adjust HINCP to constant dollars)                                                             num       housing
# 22 VEH      Vehicles (1 ton or less) available                                                                                                            chr       housing
# 23 RMSP     Number of rooms                                                                                                                               num       housing
# 24 NP       Number of persons associated with this housing record                                                                                         num       housing
# 25 GRPIP    Gross rent as a percentage of household income past 12 months                                                                                 num       housing
# 26 OCPIP    Selected monthly owner costs as a percentage of household income during the past 12 months                                                    num       housing
# 27 POVPIP   Income-to-poverty ratio recode                                                                                                                num       person 

# Vector of PUMA codes to get from CA- ACS PUMS API Call below:
puma_codes_scag <- c("02500", "03701", "03702", "03703", "03704", "03705", "03706", "03707", "03708", "03709", "03710", "03711", "03712", "03713", 
                     "03714", "03715", "03716", "03717", "03718", "03719", "03720", "03721", "03722", "03723", "03724", "03725", "03726", "03727", 
                     "03728", "03729", "03730", "03731", "03732", "03733", "03734", "03735", "03736", "03737", "03738", "03739", "03740", "03741", 
                     "03742", "03743", "03744", "03745", "03746", "03747", "03748", "03749", "03750", "03751", "03752", "03753", "03754", "03755", 
                     "03756", "03757", "03758", "03759", "03760", "03761", "03762", "03763", "03764", "03765", "03766", "03767", "03768", "03769", 
                     "05901", "05902", "05903", "05904", "05905", "05906", "05907", "05908", "05909", "05910", "05911", "05912", "05913", "05914", 
                     "05915", "05916", "05917", "05918", "06501", "06502", "06503", "06504", "06505", "06506", "06507", "06508", "06509", "06510", 
                     "06511", "06512","06513", "06514", "06515", "07101", "07102", "07103", "07104", "07105", "07106", "07107", "07108", "07109", 
                     "07110", "07111", "07112", "07113", "07114", "07115", "11101", "11102", "11103", "11104", "11105", "11106")
  
###
# Run API call to get all California ACS PUMS data from 2017-2021- 5 year ACS: CA FIPS is 06- took about 8 minutes to download API call.
ca_pums <- get_pums(
  variables = vars_list_pums,
  state = "CA", # California
  puma = puma_codes_scag,
  survey = "acs5", # acs5 is the 5 year ACS PUMS
  year = 2021, # Gets 2017-2021 ACS PUMS if acs5 set in survey option.
  recode = TRUE, # recode variable values using Census data dictionary and creates a new *_label column for each variable that is recoded.
  show_call  = TRUE # If TRUE, display call made to Census API.
)

# API CALL: Census API call: https://api.census.gov/data/2021/acs/acs5/pums?get=SERIALNO%2CSPORDER%2CWGTP%2CPWGTP%2CPUMA%2CRELSHIPP%2CSEX%2CAGEP%2CHHT%2CHISP%2CRAC1P%2CENG%2CNATIVITY%2CDIS%2CSCHL%2CESR%2CJWMNP%2CJWTRNS%2CTEN%2CKIT%2CPLM%2CWAGP%2CWKHP%2CADJINC%2CVEH%2CRMSP%2CNP%2CGRPIP%2COCPIP%2CPOVPIP&ucgid=0400000US06

# Write out to .csv for the raw data from API call to folder named "clean_data": 1826332 obs. with 49 variables
# write_csv(ca_pums, here(acs_api_fp, "ca_pums_raw.csv"))

### 
#Read in saved API call data:
ca_pums <- read_csv(here(acs_api_fp, "ca_pums_raw.csv"))

# Set cols to numeric that ended up as characters, change order of vars:
ca_pums <- ca_pums %>% mutate(across(PUMA:RAC1P, ~ as.numeric(.))) %>% 
    relocate(c("PUMA", "ST", "SPORDER", "SEX", "AGEP","HHT", "HISP","RAC1P", "ENG", "NATIVITY", "DIS", "SCHL","ESR"), .after = SERIALNO)

# Filter data to only PUMAs that we need for SCAG Region: 
# *------ PUMA 2010 -----------------------------------;
# if 		  puma  =2500 then county=25; *Imperial;
# if 3701<= puma <=3799 then county=37; *Los Angeles;
# if 5901<= puma <=5999 then county=59; *Orange;
# if 6501<= puma <=6599 then county=65; *Riverside ;
# if 7101<= puma <=7199 then county=71; *San Bernardino;
# if 11101<= puma <=11199 then county=111; *Ventura;

# SCAG County list- need to conver PUMA codes to county and then filter whole raw ACS PUMS data set to only these counties:
scag_county_list <- c("Imperial", 
                      "Los Angeles",
                      "Orange",
                      "Riverside",
                      "San Bernardino", 
                      "Ventura")
# Mutate/case_when to replicate SAS code above for creating "county" var- 
# Create county variable with SCAG counties and filter to only those rows in SCAG Counties using scag_county_list from above:
ca_pums_scag <- ca_pums %>% mutate(county = case_when(PUMA == 2500 ~ "Imperial", # 25 prefix
                                      PUMA >= 3701 & PUMA <= 3799 ~ "Los Angeles", # 37 prefix
                                      PUMA >= 5901 & PUMA <= 5999 ~ "Orange", # 59 prefix
                                      PUMA >= 6501 & PUMA <= 6599 ~ "Riverside", # 65 prefix
                                      PUMA >= 7101 & PUMA <= 7199 ~ "San Bernardino", # 71 prefix
                                      PUMA >= 11101 & PUMA <= 11199 ~ "Ventura" # 111 prefix
                                                    )) %>% 
                            filter(county %in% scag_county_list) %>% relocate("county", .after = ST) 

# Write out data filtered to only SCAG counties to folder named "clean_data": 879241 obs. with 49 variables
# write_csv(ca_pums_scag, here(acs_api_fp, "ca_pums_scag.csv"))

# # Read in SCAG county data from ca_pums_scag.csv:
ca_pums_scag <- read_csv(here(acs_api_fp, "ca_pums_scag.csv"))
# 
# # Set a list of names of the columns from API call:
# vars_ca_pums_scag <- names(ca_pums_scag)

# Remove all large datasets, leave ca_pums_scag
# rm(merged_data,merged_data_scag,housing_data,person_data,ca_pums)
```

### CA ACS PUMS Data from 2018-2022- 5 year ACS- Census API Call:
```{r 2022 ACS PUMS Data Dictionary}
library(rvest)
# Pull data dictionary from 2018-2022 ACS PUMS from website:
acs_data_dict_22 <- "https://api.census.gov/data/2022/acs/acs5/pums/variables.html" %>% 
  read_html() %>%
  html_element("table") %>% 
  html_table(convert = TRUE) %>% 
  janitor::clean_names() 

# Drop first row- was incorrectly pulled in from footer of table ("525 variables") and write to csv called :
acs_data_dict_22 <- acs_data_dict_22 %>% filter(name != "525 variables")

# write to csv called :
acs_data_dict_22 %>% write_csv(file = "data_dicts_ACS_PUMS/acs_data_dict_22.csv")

```


```{r tidycensus CA data 2022}
# List of vars to get from ACS PUMS
vars_list_pums_22 <- c("PUMA10", "PUMA20", "RELSHIPP","SEX", "AGEP", "HHT", "HISP", "RAC1P", "ENG", "NATIVITY", "DIS", "SCHL", "ESR", "JWMNP", "JWTRNS", 
                    "TEN", "KIT", "PLM", "WAGP", "WKHP", "ADJINC", "HINCP", "VEH",
                    "RMSP", "NP", "GRPIP", "OCPIP", "POVPIP")
# Check that all vars from 21 are in 22 data:
# read in acs_data_dict_22
acs_data_dict_22 <- read_csv(file = "data_dicts_ACS_PUMS/acs_data_dict_22.csv")
needed_vars_acs_data_dict_22 <- acs_data_dict_22 %>% filter(name %in% vars_list_pums_22)

# Return tibble of var_code and var_label for vars_list_pums, arrange by order of vars_list_pums:
# pums_data_desc <- pums_vars_acs5_22 %>% 
#             distinct(var_code, var_label, data_type, level) %>% 
#             filter(var_code %in% vars_list_pums) %>%
#             mutate(var_code =  factor(var_code, levels = vars_list_pums)) %>%
#             arrange(var_code)

######
###### List of vars and labels:
###### 
#   var_code  var_label                                                                                                                                   data_type   level  
#    <fct>    <chr>                                                                                                                                         <chr>     <chr>  
#  1 PUMA     Public use microdata area code (PUMA) based on 2010 Census definition (areas with population of 100,000 or more, use with ST for unique code) chr       NA     
#  2 RELSHIPP Relationship to reference person                                                                                                              chr       person 
#  3 SEX      Sex                                                                                                                                           chr       person 
#  4 AGEP     Age                                                                                                                                           num       person 
#  5 HHT      Household/family type                                                                                                                         chr       housing
#  6 HISP     Recoded detailed Hispanic origin                                                                                                              chr       person 
#  7 RAC1P    Recoded detailed race code                                                                                                                    chr       person 
#  8 ENG      Ability to speak English                                                                                                                      chr       person 
#  9 NATIVITY Nativity                                                                                                                                      chr       person 
# 10 DIS      Disability recode                                                                                                                             chr       person 
# 11 SCHL     Educational attainment                                                                                                                        chr       person 
# 12 ESR      Employment status recode                                                                                                                      chr       person 
# 13 JWMNP    Travel time to work                                                                                                                           num       person 
# 14 JWTRNS   Means of transportation to work                                                                                                               chr       person 
# 15 TEN      Tenure                                                                                                                                        chr       housing
# 16 KIT      Complete kitchen facilities                                                                                                                   chr       housing
# 17 PLM      Complete plumbing facilities                                                                                                                  chr       housing
# 18 WAGP     Wages or salary income past 12 months (use ADJINC to adjust WAGP to constant dollars)                                                         num       person 
# 19 WKHP     Usual hours worked per week past 12 months                                                                                                    num       person 
# 20 ADJINC   Adjustment factor for income and earnings dollar amounts (6 implied decimal places)                                                           chr       housing
# 21 HINCP    Household income (past 12 months, use ADJINC to adjust HINCP to constant dollars)                                                             num       housing
# 22 VEH      Vehicles (1 ton or less) available                                                                                                            chr       housing
# 23 RMSP     Number of rooms                                                                                                                               num       housing
# 24 NP       Number of persons associated with this housing record                                                                                         num       housing
# 25 GRPIP    Gross rent as a percentage of household income past 12 months                                                                                 num       housing
# 26 OCPIP    Selected monthly owner costs as a percentage of household income during the past 12 months                                                    num       housing
# 27 POVPIP   Income-to-poverty ratio recode                                                                                                                num       person 

# Vector of PUMA codes to get from CA- ACS PUMS API Call below:
puma_codes_scag <- c("02500", "03701", "03702", "03703", "03704", "03705", "03706", "03707", "03708", "03709", "03710", "03711", "03712", "03713", 
                     "03714", "03715", "03716", "03717", "03718", "03719", "03720", "03721", "03722", "03723", "03724", "03725", "03726", "03727", 
                     "03728", "03729", "03730", "03731", "03732", "03733", "03734", "03735", "03736", "03737", "03738", "03739", "03740", "03741", 
                     "03742", "03743", "03744", "03745", "03746", "03747", "03748", "03749", "03750", "03751", "03752", "03753", "03754", "03755", 
                     "03756", "03757", "03758", "03759", "03760", "03761", "03762", "03763", "03764", "03765", "03766", "03767", "03768", "03769", 
                     "05901", "05902", "05903", "05904", "05905", "05906", "05907", "05908", "05909", "05910", "05911", "05912", "05913", "05914", 
                     "05915", "05916", "05917", "05918", "06501", "06502", "06503", "06504", "06505", "06506", "06507", "06508", "06509", "06510", 
                     "06511", "06512","06513", "06514", "06515", "07101", "07102", "07103", "07104", "07105", "07106", "07107", "07108", "07109", 
                     "07110", "07111", "07112", "07113", "07114", "07115", "11101", "11102", "11103", "11104", "11105", "11106")
  
###
# Run API call to get all California ACS PUMS data from 2018-2022- 5 year ACS: CA FIPS is 06- took about 8 minutes to download API call.
ca_pums_22 <- get_pums(
  variables = vars_list_pums_22,
  state = "CA", # California
  puma = puma_codes_scag,
  survey = "acs5", # acs5 is the 5 year ACS PUMS
  year = 2022, # Gets 2018-2022 ACS PUMS if acs5 set in survey option.
  recode = TRUE, # recode variable values using Census data dictionary and creates a new *_label column for each variable that is recoded.
  show_call  = TRUE # If TRUE, display call made to Census API.
)

# API CALL: Census API call: https://api.census.gov/data/2021/acs/acs5/pums?get=SERIALNO%2CSPORDER%2CWGTP%2CPWGTP%2CPUMA%2CRELSHIPP%2CSEX%2CAGEP%2CHHT%2CHISP%2CRAC1P%2CENG%2CNATIVITY%2CDIS%2CSCHL%2CESR%2CJWMNP%2CJWTRNS%2CTEN%2CKIT%2CPLM%2CWAGP%2CWKHP%2CADJINC%2CVEH%2CRMSP%2CNP%2CGRPIP%2COCPIP%2CPOVPIP&ucgid=0400000US06

# Write out to .csv for the raw data from API call to folder named "clean_data": 1826332 obs. with 49 variables
# write_csv(ca_pums, here(acs_api_fp, "ca_pums_raw.csv"))

### 
#Read in saved API call data:
# ca_pums <- read_csv(here(acs_api_fp, "ca_pums_raw.csv"))

# Set cols to numeric that ended up as characters, change order of vars:
ca_pums <- ca_pums %>% mutate(across(PUMA:RAC1P, ~ as.numeric(.))) %>% 
    relocate(c("PUMA", "ST", "SPORDER", "SEX", "AGEP","HHT", "HISP","RAC1P", "ENG", "NATIVITY", "DIS", "SCHL","ESR"), .after = SERIALNO)

# Filter data to only PUMAs that we need for SCAG Region: 
# *------ PUMA 2010 -----------------------------------;
# if 		  puma  =2500 then county=25; *Imperial;
# if 3701<= puma <=3799 then county=37; *Los Angeles;
# if 5901<= puma <=5999 then county=59; *Orange;
# if 6501<= puma <=6599 then county=65; *Riverside ;
# if 7101<= puma <=7199 then county=71; *San Bernardino;
# if 11101<= puma <=11199 then county=111; *Ventura;

# SCAG County list- need to conver PUMA codes to county and then filter whole raw ACS PUMS data set to only these counties:
scag_county_list <- c("Imperial", 
                      "Los Angeles",
                      "Orange",
                      "Riverside",
                      "San Bernardino", 
                      "Ventura")
# Mutate/case_when to replicate SAS code above for creating "county" var- 
# Create county variable with SCAG counties and filter to only those rows in SCAG Counties using scag_county_list from above:
ca_pums_scag <- ca_pums %>% mutate(county = case_when(PUMA == 2500 ~ "Imperial", # 25 prefix
                                      PUMA >= 3701 & PUMA <= 3799 ~ "Los Angeles", # 37 prefix
                                      PUMA >= 5901 & PUMA <= 5999 ~ "Orange", # 59 prefix
                                      PUMA >= 6501 & PUMA <= 6599 ~ "Riverside", # 65 prefix
                                      PUMA >= 7101 & PUMA <= 7199 ~ "San Bernardino", # 71 prefix
                                      PUMA >= 11101 & PUMA <= 11199 ~ "Ventura" # 111 prefix
                                                    )) %>% 
                            filter(county %in% scag_county_list) %>% relocate("county", .after = ST) 

# Write out data filtered to only SCAG counties to folder named "clean_data": 879241 obs. with 49 variables
# write_csv(ca_pums_scag, here(acs_api_fp, "ca_pums_scag.csv"))

# # Read in SCAG county data from ca_pums_scag.csv:
# ca_pums_scag <- read_csv(here(acs_api_fp, "ca_pums_scag.csv"))
# 
# # Set a list of names of the columns from API call:
# vars_ca_pums_scag <- names(ca_pums_scag)

# Remove all large datasets, leave ca_pums_scag
# rm(merged_data,merged_data_scag,housing_data,person_data,ca_pums)
```


### SCAG Data from ACS PUMS:
```{r read in SCAG data}
# Read in SCAG county data from ca_pums_scag.csv:
ca_pums_scag <- read_csv(here(acs_api_fp, "ca_pums_scag.csv"))

# Set a list of names of the columns from API call:
vars_ca_pums_scag <- names(ca_pums_scag)

# Read in merged ACS data SCAG:
merged_acs_data_scag <- read_csv(here(acs_data_21_fp, "merged_acs_data_scag.csv"))
```


### Recoding variables for use in ACS PUMS Analysis: 
```{r recode}
# Race and Ethnicity: named var "race" for brevity:
# SAS example-
# *------- 1. Race_Ethnicity -------------------------------------------------;
# 
# racespan=999; 
# if hisp ge 2 then racespan=1;             	*Hispanic ;
# if hisp=1 then do;
#    if RAC1P=1 then racespan=2;              *NH White;
#    if RAC1P=2 then racespan=3;              *NH Black;
#    if 6<= RAC1P <= 7 then racespan=4;       *NH Asian & PI;
#    if 3<= RAC1P <= 5 then racespan=5;       *NH Am Indian;
#    if RAC1P >= 8 then racespan=6;           *NH Other;
#    end;
# Create "race" var based on conditions above: 
ca_pums_scag <- ca_pums_scag %>% mutate(race = case_when(HISP != 1  ~ "Hispanic/Latino",
                                                         HISP == 1 & RAC1P == 1 ~ "White", 
                                                         HISP == 1 & RAC1P == 2 ~ "Black", 
                                                         HISP == 1 & RAC1P == 6 | HISP == 1 & RAC1P == 7 ~ "Asian/Pacific Islander", 
                                                         HISP == 1 & RAC1P >= 3 & RAC1P <= 5  ~ "Native American", 
                                                         HISP == 1 & RAC1P == 8 | HISP == 1 & RAC1P == 9  ~ "Multiracial/Other"
                                                         )) 

###
# Homeownership: own
# SAS example- 
# *------- H3.Homeownership: Tenure  -------------;
# 
# *TEN Character 1
# Tenure
# b .N/A (GQ/vacant)
# 1 .Owned with mortgage or loan (include home equity loans)
# 2 .Owned free and clear
# 3 .Rented
# 4 .Occupied without payment of rent;
# 
# own=.;
# if TEN in (1 2) then own=1;
# if TEN in (3 4) then own=0;
# 
# Create "own_home" var based on conditions above: 
ca_pums_scag <- ca_pums_scag %>% mutate(own_home = case_when(TEN == 1 | TEN == 2 ~ "Homeowner",
                                                             TEN == 3 | TEN == 4 ~ "Non-Homeowner"
                                                            ))

###
# Households without a vehicle : 
# SAS example-
# *------- H1.Percent of households without a vehicle  -------------;
# *
# VEH Character 1
# Vehicles (1 ton or less) available
# b .N/A (GQ/vacant)
# 0 .No vehicles
# 1 .1 vehicle
# 2 .2 vehicles
# 3 .3 vehicles
# 4 .4 vehicles
# 5 .5 vehicles
# 6 .6 or more vehicles
# ;
# Veh_zero=0;
# if VEH=0 then Veh_zero=1
#;
#
# VEH_label 
# "No vehicles"  "1 vehicle"  "2 vehicles"  "3 vehicles"  "4 vehicles"  "5 vehicles"  "6 or more vehicles" "N/A (GQ/vacant)"

# Means of transportation to work- code- Workers who Commute by Walk, Bike, or Public Transit: named var "walk_bike_public_transport":
# SAS example-
# *JWTRNS Character 2
# Means of transportation to work
# bb .N/A (not a worker-not in the labor force, including persons .under 16 years, unemployed, employed, with a job but not at .work, Armed Forces, with a job but not at work)
# 01 .Car, truck, or van
# 02 .Bus
# 03 .Subway or elevated rail
# 04 .Long-distance train or commuter rail
# 05 .Light rail, streetcar, or trolley
# 06 .Ferryboat
# 07 .Taxicab
# 08 .Motorcycle
# 09 .Bicycle
# 10 .Walked
# 11 .Worked from home
# 12 .Other method;
# 
# Commute=.;
# if JWTRNS=1 then Commute=1;*Private;
# if JWTRNS in (2 3 4 5 6 7) then Commute=2;*Public;
# if JWTRNS in (8 9 10 12) then Commute=3;*Other;
# 
# 
# Commute2=.;
# if JWTRNS in (1 8) then Commute2=1;*Private(car or motorcycle);
# if JWTRNS in (2 3 4 5 6 7) then Commute2=2;*Public(bus, rail, taxi or ferry);
# if JWTRNS in (9 10 12) then Commute2=3;*Other(bike, walk or other);
# Create "race" var based on conditions above: 
ca_pums_scag <- ca_pums_scag %>% mutate(walk_bike_public_transport = case_when(JWTRNS %in% c(2:5,9,10)  ~ "Walk/Bike/Public",
                                                         TRUE ~ "Other Transport"
                                                         )) 
# Means of transportation to work- code- Workers who Commute by Walk, Bike, or Public Transit: named var "walk_bike_public_transport":
# SAS example-
# *------- P3.Average travel time to work (minutes)  -------------;
# 
# *JWMNP Numeric 3
# Travel time to work
# bbb .N/A (not a worker or worker who worked at home)
# 1..200 .1 to 200 minutes to get to work (Top-coded);
ca_pums_scag 


```

### Analysis and Replication of Tables and Figures for SCAG

#### Figure 5. Workers who Commute by Walk, Bike, or Public Transit by Race and Ethnicity
```{r fig 5}
### 
# Figure 5. Workers who Commute by Walk, Bike, or Public Transit by Race and Ethnicity:
# Use person level weights- PWGTP:
# walk_bike_public_transport ==  "Walk/Bike/Public"
# Get percentages based on non-home workers- drop all if JWTRNS respond == 11 (use filter not equal to 11 like this (filter(JWTRNS != 11)))
# Get Walk/Bike/Public percentages by race for all SCAG non-home workers:
scag_race_walk_bike_public <- ca_pums_scag %>% filter(JWTRNS != 11) %>% select(PWGTP, race, walk_bike_public_transport) %>%  
                                         group_by(race) %>% 
                                         summarize(
                                           total_pop = sum(PWGTP),
                                           walk_bike_public = sum(PWGTP[walk_bike_public_transport == "Walk/Bike/Public"]),  
                                           walk_bike_public_prop = walk_bike_public / total_pop, # divide by total_pop for proportion
                                           walk_bike_public_perc = round(walk_bike_public_prop*100, 2)
                                         ) %>% mutate(county = "SCAG") %>% select(county, race, walk_bike_public_perc) %>% 
                                         pivot_wider(names_from = race, values_from = walk_bike_public_perc) # make into wide table to match excel file table.

# Get Walk/Bike/Public percentages by race for all by race and by county for non-home workers:
scag_county_race_walk_bike_public <- ca_pums_scag %>% filter(JWTRNS != 11) %>% select(PWGTP, county, race, walk_bike_public_transport) %>%  
                                         group_by(county, race) %>% 
                                         summarize(
                                           total_pop = sum(PWGTP),
                                           walk_bike_public = sum(PWGTP[walk_bike_public_transport == "Walk/Bike/Public"]),  
                                           walk_bike_public_prop = walk_bike_public / total_pop, # divide by total_pop for proportion
                                           walk_bike_public_perc = round(walk_bike_public_prop*100, 2)
                                         ) %>% select(county, race, walk_bike_public_perc) %>% 
                                         pivot_wider(names_from = race, values_from = walk_bike_public_perc)

# Get no_veh percentages by county for all SCAG:
scag_county_total_walk_bike_public <- ca_pums_scag %>% filter(JWTRNS != 11) %>% 
                                         select(PWGTP, county, walk_bike_public_transport)  %>%  
                                         group_by(county) %>% 
                                         summarize(
                                           total_pop = sum(PWGTP),
                                           walk_bike_public = sum(PWGTP[walk_bike_public_transport == "Walk/Bike/Public"]),  
                                           walk_bike_public_prop = walk_bike_public / total_pop, # divide by total_pop for proportion
                                           walk_bike_public_perc = round(walk_bike_public_prop*100, 2)
                                         ) %>% select(county, walk_bike_public_perc) %>% rename(Total = walk_bike_public_perc) 
# Get no_veh percentages for all SCAG:
scag_total_walk_bike_public <- ca_pums_scag %>% filter(JWTRNS != 11) %>% select(PWGTP, walk_bike_public_transport) %>%  
                                         summarize(
                                           total_pop = sum(PWGTP),
                                           walk_bike_public = sum(PWGTP[walk_bike_public_transport == "Walk/Bike/Public"]),  
                                           walk_bike_public_prop = walk_bike_public / total_pop, # divide by total_pop for proportion
                                           walk_bike_public_perc = round(walk_bike_public_prop*100, 2)
                                         ) %>% mutate(county = "SCAG") %>% select(county, walk_bike_public_perc) %>% rename(Total = walk_bike_public_perc)

# Full join all scag_total tables:
scag_total_walk_bike_public_table <- full_join(scag_county_total_walk_bike_public,scag_total_walk_bike_public)

# Full join all scag_race tables, then left join scag_total_no_veh_table by county:
scag_walk_bike_public_table <- full_join(scag_county_race_walk_bike_public,scag_race_walk_bike_public) %>% left_join(.,scag_total_walk_bike_public_table, by = join_by(county))
scag_walk_bike_public_table

```


#### Figure 6. Figure 6. Householders without a Vehicle by Race and Ethnicity
```{r fig 6}
### 
# Figure 6. Householders without a Vehicle by Race and Ethnicity:
# Filter to only housing data(RELSHIPP == 20) and Use household weights- WGTP:
# # VEH_label 
# "No vehicles"  "1 vehicle"  "2 vehicles"  "3 vehicles"  "4 vehicles"  "5 vehicles"  "6 or more vehicles" "N/A (GQ/vacant)"
# Get no_veh percentages by race for all SCAG:
scag_race_no_veh <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, race,VEH_label) %>%  
                                         group_by(race) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           no_veh = sum(WGTP[VEH_label == "No vehicles"]),  
                                           no_veh_prop = no_veh / total_pop, # divide by total_pop for proportion
                                           no_veh_perc = round(no_veh_prop*100, 2)
                                         ) %>% mutate(county = "SCAG") %>% select(county, race, no_veh_perc) %>% 
                                         pivot_wider(names_from = race, values_from = no_veh_perc) # make into wide table to match excel file table.

# Get no_veh percentages by race and by county:
scag_county_race_no_veh <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, county, race, VEH_label) %>%  
                                         group_by(county, race) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           no_veh = sum(WGTP[VEH_label == "No vehicles"]), 
                                           no_veh_prop = no_veh / total_pop, # divide by total_pop for proportion
                                           no_veh_perc = round(no_veh_prop*100, 2)
                                         ) %>% select(county, race, no_veh_perc) %>% 
                                         pivot_wider(names_from = race, values_from = no_veh_perc)
# Get no_veh percentages by county for all SCAG:
scag_county_total_no_veh <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, county, VEH_label) %>%  
                                         group_by(county) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           no_veh = sum(WGTP[VEH_label == "No vehicles"]),
                                           no_veh_prop = no_veh / total_pop, # divide by total_pop for proportion
                                           no_veh_perc = round(no_veh_prop*100, 2)
                                         ) %>% select(county, no_veh_perc) %>% rename(Total = no_veh_perc) 
# Get no_veh percentages for all SCAG:
scag_total_no_veh <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, VEH_label) %>%  
                                         summarize(
                                           total_pop = sum(WGTP),
                                           no_veh = sum(WGTP[VEH_label == "No vehicles"]), 
                                           no_veh_prop = no_veh / total_pop, # divide by total_pop for proportion
                                           no_veh_perc = round(no_veh_prop*100, 2)
                                         ) %>% mutate(county = "SCAG") %>% select(county, no_veh_perc) %>% rename(Total = no_veh_perc)

# Full join all scag_total tables:
scag_total_no_veh_table <- full_join(scag_county_total_no_veh,scag_total_no_veh)

# Full join all scag_race tables, then left join scag_total_no_veh_table by county:
scag_no_veh_table <- full_join(scag_county_race_no_veh,scag_race_no_veh) %>% left_join(.,scag_total_no_veh_table, by = join_by(county))
scag_no_veh_table

```

#### Figure 7. Workers Commute Times (Minutes) by Mode and Race and Ethnicity
```{r fig 7}
# ### 
# # Figure 5. Workers who Commute by Walk, Bike, or Public Transit by Race and Ethnicity:
# # Use person level weights- PWGTP:
# # walk_bike_public_transport ==  "Walk/Bike/Public"
# # Get percentages based on non-home workers- drop all if JWTRNS respond == 11 (use filter not equal to 11 like this (filter(JWTRNS != 11)))
# # Get Walk/Bike/Public percentages by race for all SCAG non-home workers:
# scag_race_walk_bike_public <- ca_pums_scag %>% filter(JWTRNS != 11) %>% select(PWGTP, race, walk_bike_public_transport) %>%  
#                                          group_by(race) %>% 
#                                          summarize(
#                                            total_pop = sum(PWGTP),
#                                            walk_bike_public = sum(PWGTP[walk_bike_public_transport == "Walk/Bike/Public"]),  
#                                            walk_bike_public_prop = walk_bike_public / total_pop, # divide by total_pop for proportion
#                                            walk_bike_public_perc = round(walk_bike_public_prop*100, 2)
#                                          ) %>% mutate(county = "SCAG") %>% select(county, race, walk_bike_public_perc) %>% 
#                                          pivot_wider(names_from = race, values_from = walk_bike_public_perc) # make into wide table to match excel file table.
# 
# # Get Walk/Bike/Public percentages by race for all by race and by county for non-home workers:
# scag_county_race_walk_bike_public <- ca_pums_scag %>% filter(JWTRNS != 11) %>% select(PWGTP, county, race, walk_bike_public_transport) %>%  
#                                          group_by(county, race) %>% 
#                                          summarize(
#                                            total_pop = sum(PWGTP),
#                                            walk_bike_public = sum(PWGTP[walk_bike_public_transport == "Walk/Bike/Public"]),  
#                                            walk_bike_public_prop = walk_bike_public / total_pop, # divide by total_pop for proportion
#                                            walk_bike_public_perc = round(walk_bike_public_prop*100, 2)
#                                          ) %>% select(county, race, walk_bike_public_perc) %>% 
#                                          pivot_wider(names_from = race, values_from = walk_bike_public_perc)
# 
# # Get no_veh percentages by county for all SCAG:
# scag_county_total_walk_bike_public <- ca_pums_scag %>% filter(JWTRNS != 11) %>% 
#                                          select(PWGTP, county, walk_bike_public_transport)  %>%  
#                                          group_by(county) %>% 
#                                          summarize(
#                                            total_pop = sum(PWGTP),
#                                            walk_bike_public = sum(PWGTP[walk_bike_public_transport == "Walk/Bike/Public"]),  
#                                            walk_bike_public_prop = walk_bike_public / total_pop, # divide by total_pop for proportion
#                                            walk_bike_public_perc = round(walk_bike_public_prop*100, 2)
#                                          ) %>% select(county, walk_bike_public_perc) %>% rename(Total = walk_bike_public_perc) 
# # Get no_veh percentages for all SCAG:
# scag_total_walk_bike_public <- ca_pums_scag %>% filter(JWTRNS != 11) %>% select(PWGTP, walk_bike_public_transport) %>%  
#                                          summarize(
#                                            total_pop = sum(PWGTP),
#                                            walk_bike_public = sum(PWGTP[walk_bike_public_transport == "Walk/Bike/Public"]),  
#                                            walk_bike_public_prop = walk_bike_public / total_pop, # divide by total_pop for proportion
#                                            walk_bike_public_perc = round(walk_bike_public_prop*100, 2)
#                                          ) %>% mutate(county = "SCAG") %>% select(county, walk_bike_public_perc) %>% rename(Total = walk_bike_public_perc)
# 
# # Full join all scag_total tables:
# scag_total_walk_bike_public_table <- full_join(scag_county_total_walk_bike_public,scag_total_walk_bike_public)
# 
# # Full join all scag_race tables, then left join scag_total_no_veh_table by county:
# scag_walk_bike_public_table <- full_join(scag_county_race_walk_bike_public,scag_race_walk_bike_public) %>% left_join(.,scag_total_walk_bike_public_table, by = join_by(county))
# scag_walk_bike_public_table

```

#### Figure 22. Homeownership by Race and Ethnicity
```{r fig 22}
### 
# Figure 22. Homeownership by Race and Ethnicity:
# Filter to only housing data () and Use household weights- WGTP:
# Get homeownership percentages by race for all SCAG:
scag_race_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, race,own_home) %>%  
                                         group_by(race) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = round(ownership_prop*100, 2)
                                         ) %>% mutate(county = "SCAG") %>% select(county, race, ownership_perc) %>% 
                                         pivot_wider(names_from = race, values_from = ownership_perc) # make into wide table to match excel file table.

# Get homeownership percentages by race and by county:
scag_county_race_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, county, race,own_home) %>%  
                                         group_by(county, race) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = round(ownership_prop*100, 2) 
                                         ) %>% select(county, race, ownership_perc) %>% 
                                         pivot_wider(names_from = race, values_from = ownership_perc)
# Get homeownership percentages by county for all SCAG:
scag_county_total_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, county, own_home) %>%  
                                         group_by(county) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = round(ownership_prop*100, 2) 
                                         ) %>% select(county, ownership_perc) %>% rename(Total = ownership_perc) 
# Get homeownership percentages for all SCAG:
scag_total_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, own_home) %>%  
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = round(ownership_prop*100, 2) 
                                         ) %>% mutate(county = "SCAG") %>% select(county, ownership_perc) %>% rename(Total = ownership_perc)

# Full join all scag_total tables:
scag_total_homeownership_table <- full_join(scag_county_total_homeownership,scag_total_homeownership)

# Full join all scag_race tables, then left join scag_total_homeownership_table by county:
scag_homeownership_table <- full_join(scag_county_race_homeownership,scag_race_homeownership) %>% left_join(.,scag_total_homeownership_table)
scag_homeownership_table
```



### Write out excel file with all sheets:
-
```{r write  excel}
# Example:
# library(readxl)
# library(writexl)
# 
# sheets <- list("sheet1Name" = sheet1, "sheet2Name" = sheet2) #assume sheet1 and sheet2 are data frames
# write_xlsx(sheets, "path/to/location")
# # Read in again:
# read_xlsx("path/to/location")


# Write out table to new sheet in excel file for tabbed SCAG output, list is this format: ("name of new sheet" = df/tibble)
# Make a list that has the name or each sheets and the df/tibble that will populate it-
sheets <- list("Figure_5" = scag_walk_bike_public_table,
               "Figure_6" = scag_no_veh_table,
               "Figure_22" = scag_homeownership_table) # scag_homeownership_table is a tibble
# Pass sheet list to write_xlsx() and destination "path/filename.xlsx":
write_xlsx(sheets, "output/scag_tabbed.xlsx")

```

--- 

```{r scratch code area, eval=FALSE, include=FALSE}
# Examples of sex by county with weight by person using var "PWGTP"
ca_pums_scag %>% 
  count(county, SEX_label, wt = PWGTP)

ca_pums_scag %>% 
  mutate(ba_above = SCHL %in% c("21", "22", "23", "24")) %>% 
  group_by(county, SEX_label) %>% 
  summarize(
    total_pop = sum(PWGTP),
    mean_age = weighted.mean(AGEP, PWGTP),
    ba_above = sum(PWGTP[ba_above == TRUE & AGEP >= 25]),
    ba_above_pct = ba_above / sum(PWGTP[AGEP >= 25])
  )

ca_pums_scag %>% 
   count(county, SEX_label, wt = PWGTP)

ca_pums_scag %>% 
   count(SEX_label, wt = PWGTP)

ca_pums %>% 
   count(SEX_label, wt = PWGTP)

# Get homeownership percentages by race for all SCAG:
scag_race_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, race,own_home) %>%  
                                         group_by(race) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = scales::percent(ownership_prop, accuracy = 1)
                                         ) %>% mutate(county = "SCAG") %>% select(county, race, ownership_perc) %>% 
                                         pivot_wider(names_from = race, values_from = ownership_perc) # make into wide table to match excel file table.

# Get homeownership percentages by race and by county:
scag_county_race_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, county, race,own_home) %>%  
                                         group_by(county, race) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = scales::percent(ownership_prop, accuracy = 1) 
                                         ) %>% select(county, race, ownership_perc) %>% 
                                         pivot_wider(names_from = race, values_from = ownership_perc)
# Get homeownership percentages by county for all SCAG:
scag_county_total_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, county, own_home) %>%  
                                         group_by(county) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = scales::percent(ownership_prop, accuracy = 1) 
                                         ) %>% select(county, ownership_perc) %>% rename(Total = ownership_perc) 
# Get homeownership percentages for all SCAG:
scag_total_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, own_home) %>%  
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = scales::percent(ownership_prop, accuracy = 1) 
                                         ) %>% mutate(county = "SCAG") %>% select(county, ownership_perc) %>% rename(Total = ownership_perc)

# Full join all scag_total tables:
scag_total_homeownership_table <- full_join(scag_county_total_homeownership,scag_total_homeownership)

# Full join all scag_race tables, then left join scag_total_homeownership_table by county:
scag_homeownership_table <- full_join(scag_county_race_homeownership,scag_race_homeownership) %>% left_join(.,scag_total_homeownership_table, by = join_by(county))
scag_homeownership_table
```


---

### Replication of 
#### Figure 22. Homeownership by Race and Ethnicity
```{r new_tab, eval=FALSE}
### 
# Figure 22. Homeownership by Race and Ethnicity:
# Filter to only housing data () and Use household weights- WGTP:
# Get homeownership percentages by race for all SCAG:
scag_race_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, race,own_home) %>%  
                                         group_by(race) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = round(ownership_prop*100, 0)
                                         ) %>% mutate(county = "SCAG") %>% select(county, race, ownership_perc) %>% 
                                         pivot_wider(names_from = race, values_from = ownership_perc) # make into wide table to match excel file table.

# Get homeownership percentages by race and by county:
scag_county_race_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, county, race,own_home) %>%  
                                         group_by(county, race) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = round(ownership_prop*100, 0) 
                                         ) %>% select(county, race, ownership_perc) %>% 
                                         pivot_wider(names_from = race, values_from = ownership_perc)
# Get homeownership percentages by county for all SCAG:
scag_county_total_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, county, own_home) %>%  
                                         group_by(county) %>% 
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = round(ownership_prop*100, 0) 
                                         ) %>% select(county, ownership_perc) %>% rename(Total = ownership_perc) 
# Get homeownership percentages for all SCAG:
scag_total_homeownership <- ca_pums_scag %>% filter(RELSHIPP == 20) %>% select(WGTP, own_home) %>%  
                                         summarize(
                                           total_pop = sum(WGTP),
                                           ownership = sum(WGTP[own_home == "Homeowner"]), # sum up only those that own homes.
                                           ownership_prop = ownership / total_pop, # divide by total_pop for proportion
                                           ownership_perc = round(ownership_prop*100, 0) 
                                         ) %>% mutate(county = "SCAG") %>% select(county, ownership_perc) %>% rename(Total = ownership_perc)

# Full join all scag_total tables:
scag_total_homeownership_table <- full_join(scag_county_total_homeownership,scag_total_homeownership)

# Full join all scag_race tables, then left join scag_total_homeownership_table by county:
scag_homeownership_table <- full_join(scag_county_race_homeownership,scag_race_homeownership) %>% left_join(.,scag_total_homeownership_table)
scag_homeownership_table
```
